---
title: "Evaluate accuracy"
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
# description: |
#   An overview of the nowcastr package.
vignette: >
  %\VignetteIndexEntry{Evaluate accuracy}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---



To evaluate model accuracy, `nowcast_eval()` performs historical backtesting. 
It iteratively applies `nowcast_cl()` to past dates, censoring data that would have been unavailable at each point in time. By comparing these predicted values against the observed values (the raw data available at the time), we can quantify the model's added value.

We use two indicators:

- **Win Rate**: The frequency with which the model's absolute error is lower than the observation's absolute error. A value **>50%** suggests the nowcast is more reliable than the raw data.

- **Differential sMAPE ($\Delta\text{sMAPE}$)**:

  $$\Delta\text{sMAPE} = \text{sMAPE}_{\text{obs}} - \text{sMAPE}_{\text{pred}}$$

  This measures the average reduction in symmetric error. A **positive value** indicates the model improves accuracy over the initial report, while a **negative value** suggests the raw data was already more accurate.



## Run evaluation

You can run the evaluation with all the same parameters as `nowcast_cl()`.  
`nowcast_eval()` has only one additional parameter: `n_past`, which controls how many steps in the past you wish to run a nowcast on.


```{r}
library(nowcastr)
nc_eval_obj <-
  nowcast_demo %>%
  nowcast_eval(
    n_past = 10,
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    time_units = "weeks",
    do_model_fitting = TRUE
  )
```

This will return an S7 object with 2 slots for 2 datasets.  

- `nc_eval_obj@detail` contains detailed results, and  
- `nc_eval_obj@summary` is summarised by `group_cols` and `delay`.

```{r}
nc_eval_obj@detail
nc_eval_obj@summary
```


## Plots

### Plot aggregated indicators

```{r, fig.width=9, fig.asp=5.5/10, out.width="100%"}
#| warning: false
plot_nowcast_eval(nc_eval_obj, delay = 0)
```

### Plot one indicator by delay

```{r, fig.width=9, fig.asp=5.5/10, out.width="100%"}
#| warning: false
plot_nowcast_eval_by_delay(nc_eval_obj, indicator = "smape_diff_med")
```

### Plot raw values, for one delay
- predicted values
- observed values (i.e. reported at the time)
- last reported values (ground truth)

```{r, fig.width=9, fig.asp=5.5/10, out.width="100%"}
#| warning: false
plot_nowcast_eval_detail(nc_eval_obj, delay = 0)
```




## Evaluate Scenarios

We can test if accuracy of nowcasts improve with or without `fill_future_reported_values()`:

```{r, fig.width=9, fig.asp=5.5/10, out.width="100%"}
#| warning: false

nc_eval_obj_with_fill <-
  nowcast_demo %>%
  fill_future_reported_values(
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    max_delay = "auto"
  ) %>%
  nowcast_eval(
    n_past = 10,
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    time_units = "weeks",
    do_model_fitting = TRUE
  )
```


```{r, fig.width=9, fig.asp=5.5/10, out.width="100%"}
#| warning: false

library(dplyr)
indicator <- "smape_diff_med"

scenario_a <- nc_eval_obj@summary %>%
  dplyr::select("group", "delay", "smape_diff_med") %>%
  dplyr::mutate(scenario = "A = no fill")

scenario_b <- nc_eval_obj_with_fill@summary %>%
  dplyr::select("group", "delay", "smape_diff_med") %>%
  dplyr::mutate(scenario = "B = fill")

## quick mean of everything
dplyr::bind_rows(scenario_a, scenario_b) %>%
  dplyr::filter(delay <= 3) %>% ## predictions for older data are not that interesting
  dplyr::summarise(
    .by = c(scenario),
    avg_smape_diff_med = mean(smape_diff_med, na.rm = T)
  ) %>%
  dplyr::mutate(tag = dplyr::if_else(avg_smape_diff_med == max(avg_smape_diff_med), "better", "worse")) %>%
  print()
```


```{r, fig.width=9, fig.asp=5.5/10, out.width="100%"}
#| warning: false

## Comparison plot
library(ggplot2)
dplyr::bind_rows(scenario_a, scenario_b) %>%
  dplyr::summarise(
    .by = c(delay, scenario, group),
    smape_diff_med = mean(smape_diff_med, na.rm = T)
  ) %>%
  ggplot(aes(x = delay, y = smape_diff_med, color = scenario)) +
  geom_point() +
  geom_line() +
  facet_wrap(~group) +
  # ggplot2::scale_y_continuous(labels = scales::label_percent()) +
  theme_nowcastr() +
  labs(
    y = "Differential sMAPE",
    x = "Delay",
    color = "Scenario",
    title = "Compare Differential sMAPE of 2 scenarios",
    subtitle = "Higher is better"
  )
```