---
title: "Evaluate accuracy"
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
# description: |
#   An overview of the nowcastr package.
vignette: >
  %\VignetteIndexEntry{Evaluate accuracy}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

To evaluate the accuracy of the model, we can measure whether the nowcasted 
values outperformed the reported values in the past. `nowcast_eval()` does 
this by running `nowcast_cl()` multiple times on past data, each time 
truncating the knowledge horizon so that only data available at that point 
in time is used.


## Run evaluation

You can run the evaluation with all the same parameters as `nowcast_cl()`.  
`nowcast_eval()` has only one additional parameter: `n_past`, which controls how many steps in the past you wish to run a nowcast on.


```{r}
library(nowcastr)
nc_eval_obj <-
  nowcast_demo %>%
  nowcast_eval(
    n_past = 10,
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    time_units = "weeks",
    do_model_fitting = TRUE
  )
```

This will return an S7 object with 2 slots:

```{r}
nc_eval_obj@detail %>% dplyr::glimpse(0)
```


```{r}
nc_eval_obj@summary %>% dplyr::glimpse(0)
```


## Plots

### Plot aggregated indicators

- "SMAPE Improvement median" = median of the difference between SMAPE(observed) and SMAPE(predicted)
- "Proportion Better = proportion of predictions that outperform base values (-50% to center around zero)

```{r}
#| warning: false
plot_nowcast_eval(nc_eval_obj, delay = 0)
```

### Plot one indicator by delay / for one indicator

```{r}
#| warning: false
plot_nowcast_eval_by_delay(nc_eval_obj, indicator = "SMAPE_improvement_med")
```

### Plot raw values / for one delay
- predicted values
- base reported values, at the time
- last reported values

```{r}
plot_nowcast_eval_detail(nc_eval_obj, delay = 0)
```




## Evaluate Scenarios

We test if accuracy of nowcasts improve with `fill_future_reported_values()`:

```{r}
nc_eval_obj_with_fill <-
  nowcast_demo %>%
  fill_future_reported_values(
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    max_delay = "auto"
  ) %>%
  nowcast_eval(
    n_past = 10,
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    time_units = "weeks",
    do_model_fitting = TRUE
  )
```

```{r}
#| warning: false
plot_nowcast_eval(nc_eval_obj_with_fill, delay = 0)
```