---
title: "Evaluate accuracy"
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
# description: |
#   An overview of the nowcastr package.
vignette: >
  %\VignetteIndexEntry{Evaluate accuracy}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

To evaluate the accuracy of the model, we can measure whether the nowcasted 
values outperformed the reported values in the past. `nowcast_eval()` does 
this by running `nowcast_cl()` multiple times on past data, each time 
truncating the knowledge horizon so that only data available at that point 
in time is used.


## Run evaluation

You can run the evaluation with all the same parameters as `nowcast_cl()`.  
`nowcast_eval()` has only one additional parameter: `n_past`, which controls how many steps in the past you wish to run a nowcast on.


```{r}
library(nowcastr)
nc_eval_obj <-
  nowcast_demo %>%
  nowcast_eval(
    n_past = 10,
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    time_units = "weeks",
    do_model_fitting = TRUE
  )
```

This will return an S7 object with 2 slots for 2 datasets.  

- `nc_eval_obj@detail` contains detailed results, and  
- `nc_eval_obj@summary` is summarised by `group_cols` and `delay`.

```{r}
nc_eval_obj@detail
nc_eval_obj@summary
```


## Plots

### Plot aggregated indicators

- "SMAPE Improvement median" = median of the difference between SMAPE(observed) and SMAPE(predicted)
- "Proportion Better = proportion of predictions that outperform base values (-50% to center around zero)


```{r, fig.width=9, fig.asp=5.5/10, out.width="100%"}
#| warning: false
plot_nowcast_eval(nc_eval_obj, delay = 0)
```

### Plot one indicator by delay / for one indicator

```{r, fig.width=9, fig.asp=5.5/10, out.width="100%"}
#| warning: false
plot_nowcast_eval_by_delay(nc_eval_obj, indicator = "SMAPE_improvement_med")
```

### Plot raw values / for one delay
- predicted values
- base reported values, at the time
- last reported values

```{r, fig.width=9, fig.asp=5.5/10, out.width="100%"}
#| warning: false
plot_nowcast_eval_detail(nc_eval_obj, delay = 0)
```




## Evaluate Scenarios

We can test if accuracy of nowcasts improve with or without `fill_future_reported_values()`:

```{r, fig.width=9, fig.asp=5.5/10, out.width="100%"}
#| warning: false

nc_eval_obj_with_fill <-
  nowcast_demo %>%
  fill_future_reported_values(
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    max_delay = "auto"
  ) %>%
  nowcast_eval(
    n_past = 10,
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    time_units = "weeks",
    do_model_fitting = TRUE
  )

# plot_nowcast_eval(nc_eval_obj_with_fill, delay = 0)
```


```{r}
#| warning: false

library(dplyr)

# indicator <- "proportion_pred_is_better"
indicator <- "SMAPE_improvement_mean"

scenario_a <- nc_eval_obj@summary %>%
  dplyr::select("group", "delay", ind = all_of(indicator)) %>%
  mutate(scenario = "A = no fill")

scenario_b <- nc_eval_obj_with_fill@summary %>%
  dplyr::select("group", "delay", ind = all_of(indicator)) %>%
  mutate(scenario = "B = fill")

## quick mean of everything
dplyr::bind_rows(scenario_a, scenario_b) %>%
  filter(delay <= 3) %>% ## predictions for older data are not that interesting
  dplyr::summarise(
    .by = c(scenario),
    ind = mean(ind, na.rm = T)
  ) %>%
  mutate(tag = if_else(ind == max(ind), "better", "worse")) %>%
  print()
```


```{r}
#| eval: false


## Example comparison plot
library(ggplot2)
dplyr::bind_rows(scenario_a, scenario_b) %>%
  dplyr::summarise(
    .by = c(delay, scenario, group),
    ind = mean(ind, na.rm = T)
  ) %>%
  ggplot(aes(x = delay, y = ind, color = scenario)) +
  geom_point() +
  geom_line() +
  facet_wrap(~group) +
  # ggplot2::scale_y_continuous(
  #   labels = scales::label_percent()
  # ) +
  theme_nowcastr(20) +
  labs(
    y = "Mean SMAPE improvement", x = "Delay",
    color = "Scenario",
    title = "Compare accuracy of 2 scenarios",
    subtitle = "Higher is better"
  )
```