---
title: "Introduction to nowcastr"
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
# description: |
#   An overview of the nowcastr package.
vignette: >
  %\VignetteIndexEntry{Introduction to nowcastr}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(dplyr) # Ensure pipe operator is available
```

Nowcasting is the process of estimating the current state of a phenomenon when the data are incomplete due to reporting delays. The **nowcastr** package implements the chain-ladder method for nowcasting, supporting both non-cumulative delay-based estimation and model-based completeness fitting (*e.g.*, logistic or Gompertz curves). This vignette provides a quick start guide to using the package with demo data.

## Setup

The package is available on GitHub. Install it with:

```{r}
#| eval: false
devtools::install_github("whocov/nowcastr")
```
```{r}
library(nowcastr)
```



## Data Structure

Your dataset must contain at least three columns:

- **occurrence date**: when the event happened
- **reporting date**: when the event was reported
- **value**: the observed count/value
- \<*groups*\>: none, one or multiple grouping columns: *e.g.* `group_cols = c("group") # or c("region", "disease")`

The package includes a demo dataset `nowcast_demo` that follows this structure:

```{r}
head(nowcast_demo, 10)
```

The demo data also includes a `group` column for demonstrating grouped processing.
The dataset can also have multiple group-by columns for batch processing.



## Workflow

A typical nowcasting workflow with **nowcastr** involves the following steps.



### 1. Visualize Input Data

Before nowcasting, inspect the reporting pattern of your data:

```{r}
nowcast_demo %>%
  plot_nc_input(
    option = "triangle",
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group"
  )
```

The "millipede" plot provides an alternative view of delays:

```{r}
nowcast_demo %>%
  plot_nc_input(
    option = "millipede",
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group"
  )
```



### 2. Prepare Data (Optional)

If your data contains future occurrence dates with no reports yet, you may want to fill these with the last known reporting values to ensure consistent time units:

```{r}
data_filled <- nowcast_demo %>%
  fill_future_reported_values(
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    max_delay = "auto"
  )
data_filled %>%
  plot_nc_input(
    option = "triangle",
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group"
  )
```

This step is optional; `nowcast_cl` can handle unfilled data.



### 3. Run Nowcast

Perform the nowcasting using the chain-ladder method:

```{r}
nc_obj <-
  data_filled %>%
  nowcast_cl(
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = "group",
    time_units = "weeks",
    do_model_fitting = TRUE
  )
```

The `nowcast_cl()` function returns a `nowcast_results` object containing predictions, delay distributions, completeness estimates, and parameters.



### 4. Explore Results

Access the components of the result object:

```{r slots}
nc_obj@results # Final nowcasted values
nc_obj@delays # Delay distribution
nc_obj@completeness # Data with completeness estimates
str(nc_obj@params) # Parameters used
```

Visualize the results:

```{r plots}
#| warning: false
plot(nc_obj, which = "delays") # Delay distribution
plot(nc_obj, which = "results") # Nowcast time series
```





## How It Works

The chain-ladder method estimates "completeness" for each delay bucket:

- **Delay** = reporting date - occurrence date
- **Completeness** = observed value / last reported value (approximation of true value)
- **Average completeness** per delay bucket (across occurrence dates)
- **Nowcast** = observed value / average completeness

Recent occurrence dates have shorter delays and lower completeness. The method upweights these observations to estimate the true count.





### Grouped Processing

You can nowcast multiple groups (e.g., regions, diseases) in a single call by specifying multiple grouping columns:

```{r grouped}
#| eval: false

nowcast_cl(
  # ...
  group_cols = c("region", "disease")
)
```





## Other Utility Functions


### Calculate Retro Scores of input data

retro_score = number of actual value changes / max possible value changes

```{r calculate_retro_score}
# Calculate retro-scores (= number of actual value changes / max possible value changes)
nowcast_demo %>%
  calculate_retro_score(
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = c("group")
  )
```


### Remove duplicated data
This can be useful to reduce size without losing information.

```{r rm_repeated_values}
# Remove duplicate reported values (same value and higher reporting date)
nowcast_demo %>%
  rm_repeated_values(
    col_date_occurrence = date_occurrence,
    col_date_reporting = date_report,
    col_value = value,
    group_cols = c("group")
  )
```

